Large-Scale Marine Protected Areas in the World's Largest Tuna Fishery

### 2019-02-19 docs/PIPA_Palau.tex
Added new figures and updated MS

% Keywords are not mandatory, but authors are strongly encouraged to provide them. If provided, please include two to five keywords, separated by the pipe symbol, e.g:
\keywords{Marine Protected Areas $|$ Fisheries $|$ Purse Seining $|$ Marine Conservation}

\begin{abstract}
The Parties to the Nauru Agreement (PNA) is a system for managing the world’s largest skipjack tuna fishery, covering an area of 14.6 million km\textsuperscript{2} in the Pacific Ocean. In 2015, PNA member Kiribati implemented one of the world's largest conservation areas: The Phoenix Island Protected Area (397,447 km\textsuperscript{2}). It is still unclear what the effects of such a big MPA have been and what effects are likely to occur in the future. We use identification of fishing activity via Automatic Identification Systems (AIS) and causal inference techniques evaluate the behavioral changes and spatial redistribution of tuna purse seiners due to the implementation of a Large-Scale Marine Protected Area. Our work provides three main findings: 1) we observe a crowding effect for the first months after implementation, 2) aggregate fishing effort remains relatively constant; and 3) vessels that fished inside the protected area redistribute to adjacent waters, including the high seas, 4) the redistribution of vessels results in a reduction of 2500 vessel-days into Kiribati, which represents a loss of ~30 million USD. This suggests that rights to fish in PNA waters might now be less valuable. Unfortunately, without detailed data on vessel catch, we cannot conclude whether vessels are better or worse off financially following the closure. We use our results to inform predictions of the impacts of a proposed LSMPA in Palau (a PNA member) and estimate potential losses to range from \$2.5 to \$11 million annually. Our results not only provide an impact evaluation of the effect of LSMPAs on fishing activity, but provide insights into vessel redistribution dynamics, which may have large ecological and economic implications for marine conservation and fisheries management in developing countries. As countries continue to implement LSMPAs as a way to reach ocean protection targets, managers should consider how fishing effort will change in space and time to ensure that fishing effort is not just displaced elsewhere.
\end{abstract}

@@

\dropcap{H}umans are increasingly utilizing the oceans. Conservation and development targets are likely to cause activities such as off-shore aquaculture, conservation, deep-sea mining and and fisheries to compete for space. As we move forward in the blue economy, we must understand the potential effects of displacing one activity over the other. For example, global international goals aim to protect 10\% of the ocean environment by 2020. In an effort to meet this target, there has been a rapid increase in MPA coverage \citep{wood_2008,sala_2018}, largely driven by a small number of Large-Scale Marine Protected Areas (LSMPAs), some of which ocurr in the world's largest fishing grounds for tuna. While today a small number of LSMPAs make up at least 80\% of the managed areas in the ocean \citep{toonen_2013}, very little is known about their human dimensions and implication for fisheries \citep{gray_2017}. One issue of particular importance is that of the displacement or redistribution of fishing effort, which may influence the outcomes of a spatial closure \citep{smith_2003}.

One of the most notable Large-Scale Marine Protected Areas recently implemented is the Phoenix Island Protected Area (PIPA) in Kiribati. The area implemented in January 1st of 2015 has an extension of 397,447 km\textsuperscript{2} and represents was implemented within an area that produces 50\% of the World's tuna. Tuna purse seine fisheries in the region are collectively managed under a Vessel-day scheme (VDS) by nine countries commonly refered to as the Parties to the Nauru Agreement (PNA). Members include the Federated States of Micronesia, Kiribati, the Marshall Islands, Nauru, Palau, Papua New Guinea, the Solomon Islands, and Tuvalu. Tokelau joined the PNA group in 2012 and started selling access rights in 2013. The Nauru Agreement regulates access of foreign vessels (\emph{i.e.} those from non-PNA countries). Holding 80\% (14.6 million km\textsuperscript{2}) of historical skipjack tuna purse seine grounds within their Exclusive Economic Zones (EEZ), PNA countries have achieved greater bargaining power when providing fishing access to foreign fleets \cite{havice_2010}. The revenue from access fees represents up to 50\% of government revenue from some of this countries.

LSMPAs were initially assumed to have little social implications due to their general remoteness. However, there have been calls to incorporate the human dimensions into LSMPA management and evaluation \citep{agardy_2011,gray_2017}. Most research incorporating these dimensions has focused on governance and enforcement of LSMPAs \cite{alger_2017,christie_2017}, but they are yet to be the focus of economic analyses \citep{gray_2017}. For example, the anticipation of PIPA lead to preemptive overfishing, which will likely erod or delay the expected benefits of the intervention \cite{mcdermott_2018}. However, a reduction in fishing area without proper adjustments to the total vessel-days permited every year is likely to increase fishing intensity in the remaining open waters. While no studies have assessed the immplications of PIPA, other PNA members have pledged the implementation of LSMPAs by 2020.

@@

\begin{figure}
\centering
\includegraphics{img/sp_corr.pdf}
\caption{\label{fig:sp_corr}Number of cells that had treated and control vessels (A) and spatial correlation in the presence-absence of treated and control vessels per cell (B). The solid lines represent the 4\textsuperscript{th} degree polynomial fit reported in \ref{tab:sp_corr}}
\end{figure}

@@

The crowding effect combined with the reduction of hours spent in Kiribati and PNA waters overall suggests that treated vessels have redistributed elsewhere, meaning that they no longer buy access fees from PNA countries. To quantify the potential impacts of this leakage, we estimate the total annual vessel-days received by all PNA countries (Fig. \ref{fig:all_PS_VDS_year}). In this case we look at all 313 vessels, but continue to group them as treated and control vessels, as well as other vessels that we do not observe continuously or came into the fishery after the implementation of PIPA. We term this last group as ``others''. We find that the treated vessels spent 3,600 less days in PNA waters when comparing 2016 and 2017, while control and other vessels increased their PNA fishing by 411 and 1687 for the same period, respectively (Figs. \ref{fig:all_PS_VDS_year}A and \ref{fig:included_PS_VDS_year_DiD}). Furthermore, in 2017 treated vessels exhibit the lowest allocation of hours to PNA waters at just 10,026 vessel-days. This is lower than pre-implementation levels of 2012 - 2013 (mean $\pm$ sd of 12,552.46 $\pm$ 1396.46; we exclude 2014 due to the blue paradox effect \cite{mcdermott_2018}). The 3,600 vessel-day reduction by treated between 2016 and 2017 vessels corresponds to \$43.2 million USD at an average vessel-day price of \$12,000 USD. The 2,526 vessel-day reduction would then represent forgone profits in the order of 30.3 million USD. %We need to come up with an identification strategy for this, I believe. I simple DID doesn't work, because the post-trend would assume that there is no upper bound to effort (current DiD estimate of these data suggest a 8,700 vessel-days reduction; or 7,400 vessel-days if we account for the blue paradox with some fixed effects).

Looking at the total anual vessel-days allocated by all vessels to all countries, we see that the largest reductions ocurr for Kiribati, while Papua New Guinea exhibits a proportional increase (Fig. \ref{fig:all_PS_VDS_year}B).

From the government side, we combine annual license revenue and EEZ-specific catch statistics compiled from the Pacific Islands Forum Fisheries Agency (FFA\footnote{https://www.ffa.int/node/2050}).

\begin{figure}
\centering
	\includegraphics{img/all_PS_VDS_year.pdf}
	\caption{\label{fig:all_PS_VDS_year}A)Observed vessel-days for all countries by treated, control, and excluded vessels.}
\end{figure}

\begin{figure}
\centering
	\includegraphics{img/included_PS_VDS_year_DiD.pdf}
	\caption{\label{fig:included_PS_VDS_year_DiD}Vessel days spent inside A) PNA waters and B) Kiribati waters by group of vessels.}
\end{figure}

\subsection{Potential Revenue Loss Palau}

@@

\begin{figure}
\centering
	\includegraphics{img/financial_indicators.pdf}
	\caption{\label{fig:financial_indicators}Financial indicators for PNA countries. A) Revenue from fishing licenses fees, B) Catches in each countrie's EEZ. Vertical dashed line denotes implementation of PIPA.}
\end{figure}

\clearpage
\begin{figure}
\centering
\includegraphics{img/all_panels.pdf}
\caption{\label{fig:all_panels}Time series showing monthly averages for eight variables of interest: A) Fishing hours, B) Non-fishing hours at-sea, C) Proportion of fishing hours to total hours at-sea, D) Distance traveled, E) Mean distance from port for fishing evetns, F) Mean distance from shore for fishing events, I) Monthly hours spent in Kiribati waters, J) Proportion of fishing hours spent in PNA waters.}
\end{figure}
\clearpage

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

The pattern observed here could just be an artifact
of satellite detections increasing though time. Therefore we also
calculate a spatial correlation of presence/absence and observe similar
patterns (Fig. \ref{fig:sp_corr}B). Spatial overlap increases
in 2015 as vessels from PIPA are excluded, but then starts to decline.
Our 4th degree polynomial shows the expected patterh of a relative
constant trend before implementation which is followed by rapid increased
and subsequent decrease.

Recent advances in satellite tracking technologies and near real-time
identification of fishing activity provide us with an opportunity to ask: how do fishers respond to the implementation of an LSMPA? Does effort increase or decrease and where does it go? We use identification of fishing activity via Automatic Identification Systems (AIS) and causal inference techniques to describe
the behavioral changes and spatial redistribution of the industrial tuna
purse seine fleet due to the implementation of the Phoenix Islands Protected Area. We use the same data to hypothesize what might be the impacts of the proposed Palau National Marine Sanctuary. These are two of the largest protected areas on the planet and both are within the Parties to the Nauru Agreement (PNA) region, which is a group of (originally eight and now) nine Pacific Islands countries that collectively manage their valuable tuna stocks.

Our work is novel in the sense that it provides empirical evidence of
the effect of Large Scale Marine Protected Areas on fishing behavior and
distribution and can help guide future interventions. Understanding how
effort is displaced from LSMPAs might provide insights into how distant water fishing fleets would react to a high seas closure.

The existing empirical literature focuses on small MPAs and has been criticized for confounding correlation with causation \citep{ferraro2018causal}.

The paper is organized as follows: Section \ref{background} provides more
information on Large Scale Marine Protected Areas and some background on
empirical studies of effort redistribution. An overview of the Nauru
Agreement and associated countries, a description of the fleet that
operates in the region, and a brief history of PIPA and PNMS is also included.
Section \ref{methods} describes our data and identification strategy.
Section \ref{results} presents our results, and Section \ref{discussion}
discusses our
results.

\section{Background}\label{background}

@@

\subsection{Nauru agreement and the Phoenix Islands Protected Area}
\label{nauru-agreement-and-the-phoenix-island-protected-area}

@@

\section{Methods}\label{methods}

\subsection{Data}\label{data}

Automatic Identification Systems (AIS) are on-board devices that provide
at-sea safety and prevent ship collisions by broadcasting vessel
position, course, and activity to surrounding vessels. These broadcast
messages can be received by satellites and land-based antennas. GFW then uses 
machine learning techniques (convolutional neural networks) on the broadcast messages
to infer what type of fishing is taking place and where it is taking place, thus
allowing the estimation of near real-time fishing events globally \citep{kroodsma_2018}.

Our analysis focuses on tune purse seine vessels, the most important fishery
for PNA
countries.\footnote{We perform some of the same analyses for longliners and include them in the Appendix}
We identify a total of 186 purse seiners that fished in PNA waters at
least once before 2015. Vessels that entered the fishery after 2015 were not
exposed to the policy intervention in the pre-treatment period and are
therefore excluded from our analyses, leaving us with 125 vessels.
We identify 90 vessels that have fished inside PIPA
at least once since 2012; 72 did so before the announcement
of the closure (\emph{i.e.} 09/01/2014 \emph{sensu} \citep{mcdermott_2018}).
From these, we only observe before and after data for 64 tuna purse seiners.
On the other hand, 45 vessels never fished inside PIPA during the 2012 - 2015 period, 
but we only observe befora and after data for 28 of these.

\subsection{Analyses}\label{analyses}

Our first analysis focuses on identifying the response of 
vessels to the PIPA closure. We use daily fishing and non-fishing hours,
daily proportion of fishing vs. non-fishing hours, daily distance traveled (km), 
distance from shore (km) and distance from home port(km) for fishing and non-fishing events,
and proportion of total fishing hours allocated to Kiribati waters and PNA waters
as our main outcomes of interest.
We compare these outcomes before and after the implementation
of PIPA using a Difference-in-Differences approach. Our main
specification is the following:

where $log(y_{i,t})$ is the log-transformed outcome of interest for vessel $i$ on day $t$.
A dummy variable $Post_t$ takes the value of 0 for all
dates prior to PIPA implementation and a value of 1 for all dates
following PIPA implementation. $Treat_i$ is a dummy
variable indicating whether a vessel belongs to the treatment ($Treat_i = 1$) or control
($Treat_i = 0$) group. $\alpha$ is
the standard intercept term, $beta_1$ captures the temporal trend,
$beta_2$ captures the initial difference between treated and control groups,
and $beta_3$ is our parameter of interest: the Difference-in-Differences estimate capturing
the treatment effect. Finally, $phi_t$ and $\gamma_i$ represent
month and flag dummies that account for seasonality or
country-level management interventions. We test for other more complex specifications
that interact a quarterly dummy or year-month dummies with the treatment group
and find qualitatively the same results\footnote{I actually need more time to run these, but I don't think they'll change}.

Our second part of the analysis focuses on the redistribution of fishing
effort, where institutions may play an important role. As stated
before, LSMPAs often span the entirety of an EEZ. In the case of purse
seining in the PNA, vessels purchase access to the fishery at the
beginning of the season. If a vessel decides to fish within the PNA,
they've already made the decision to pay for access, expecting higher
returns from fishing there than in the high seas. In the particular case
of PIPA, one would expect that a vessel previously holding a permit to fish in
Kiribati waters would 
most likely continue to purchase access to PNA waters (either in Kiribati's
remaining open areas or a nearby EEZ). Alternatively, if a vessel fishes illegally
in Kiribati waters before the implementation, one would expect them to
reallocate to other regions with equal or less enforcement. In this
case, they might then choose to move to other Kiribati waters, waters
of countries that have similar enforcement levels, or to the high seas.

We discretize spatial units by creating a polygon for
PIPA and distinct
spatial units for each geographically separate EEZ of each country. Some vessels might shift
from EEZs into the high seas, but we are interested in knowing
\emph{where} in the high seas, so we incorporate additional regions by
using a 1-degree buffer of the high seas\footnote{Currently running the 5-degree version too}
The rest of the high seas are merged into a single spatial unit. For
example, if we were to do this only for Kiribati, we would have 8
spatial units: PIPA, three EEZ units, three 1-degree buffers of high seas  around each of the EEZ regions.
around each EEZ, and the rest of the high seas. We clipped any overlapping
high seas buffers to avoid duplication.

We then take these spatial polygons and rasterize them to a 1-degree grid
(see Figure \ref{fig:raster_rgn} in the Appendix), which we then use to rasterize points
of fishing activity. For each cell, we calculate the total monthly
fishing hours by treated and control vessels (Fig. \ref{fig:fishing_raster}).
Since we are interested in the redistribution
of fishing effort, we use this gridded data to calculate the proportion
of fishing hours that each region receives each month relative to the
total fishing hours observed across all regions and all cells in that
month.

We use this measure to evaluate the change in effort allocation by
regressing it on the interaction between a year dummy and a dummy for
regions:

\begin{figure*}[h]
\begin{align*}
y_{i,t} = \alpha + \beta_1Year_t + \beta_{2,i}Region_i + \beta_{3,i}Year_t \times Region_i+ \epsilon_{i,t}
\end{align*}
\end{figure*}

Our variable of interest, $y_{i,t}$ represents the proportion of
fishing hours that region $i$ receives at time $t$. Years are
modeled as factors ($Year_t$), using 2012 as the reference level.
$Region$ is a dummy variable for regions defined above. Our vector of parameters of
of interest is $\beta_{3,i}$, which captures the yearly by-country
change in proportional allocation of fishing effort relative to 2012.

Finally, we inspect the crowding effects that may arise due to the net reduction
in fishing area. We use the rasterized fishing effort and calculate two
potential indices of crowding. For the first one, we count the number of cells
that had fishing activity from treated and control vessels over a given month. For the second one,
we calculate the correlation of presence/absence of fishing events between both groups
over one month. If the implementation of the reserve induces a crowding effect,
we would expect to observe no trend before the implementation of the reserve.
The implementation would lead to an increase in these measures, which, 
if vessels avoid crowding, should then come back down. We therefore expect to
have three inflection points: 1) in the transition between no trend and initial crowding due to MPA
implementation, 2) When the crowding has reached its maximum and starts to decrease,
and 3) when the decrease levels off, pressumably to pre-MPA levels.
For this reason, we fit a 4th degree polynomial to our monthly indices. We do so by
centering our time series of crowding indices on the day of implementation. Our
explanatory variable is therefore the number of months before or after the implementation.
For example, since PIPA was implemented in January 1st of 2015, December of 2014 
has a value of -1, and Feb of 2015 would receive a value of 1.

\section{Results2}

Regression coefficients for our DiD analysis are shown in Table
\ref{tab:main_DID}. Our DiD analysis shows that PIPA closure
had no effect on the number of fishing hours, but that it caused
non-fishing at-sea hours to increase by 1\%, This coefficient is
only marginally significant. The proportion of
fishing to non-fishing hours shows a positive change, but it is not
significant. However, treated vessels travel on average 19\% less
than control vessels, and opperate closer to shore and port.
Furthermore, PIPA implementation caused fishing events of treated
vessel to occur 23\% and 4\% closer to port and to shore, respectively.
Other specificationf that exclude fixed effects produce similar estimates (Fig. \ref{fig:other_specifications})
Since vessels don't seem to be fishing a lot less, an obvious question arises: where are they
fishing now?

Figure \ref{fig:fishing_raster} provides a spatial representation of
these changes. Note that the 2013 effort distribution is roughly similar
across vessels. Then, in 2014, there is a sharp increase in fishing hours
by treated vessels inside PIPA. In 2015, treated vessels
then allocate more effort to the easternmost Kiribati EEZ and the high seas.
In 2016, the spatial distribution of
fishing effort is again similar across groups. It is evident that the
increase in relative fishing effort is greater for regions closer to
PIPA.

Along a similar line, Figure \ref{fig:redist_trend} shows a detailed
temporal evolution of the relative allocation of fishing
effort.  It shows the change in fishing effort inside PIPA, including the
preemptive fishing and immediate reduction previously reported
\citep{mcdermott_2018}. Recall that to evaluate the redistribution of fishing effort
we only track fishing vessels that belong to the treated group.
Kiribati waters show an increase in fishing effort for 2015;
the same pattern is observed for their respective High Seas buffers,
and the general High Seas. This suggests fishing effort inside PIPA has moved to the
high seas around Kiribati instead of paying to fish within alternative Kiribati EEZ waters.

We then move on to our regression results, presented as a figure showing
the year-region interaction coefficient estimates for each region as a
stream of changes through time (Fig. \ref{fig:mean_change} and Table
\ref{tab:mean_change}). As noted before, the marginal change in the
relative allocation of fishing effort relative to 2012 increases in 2014
for PIPA, followed by an abrupt reduction in 2015 and afterwards. This 
reduction confirms that effort within PIPA has been virtually eliminated.
Most coefficients are not statistically significant, but
follow the same trends previously discussed, with a tendency for regions
closer to PIPA to show an increase in the relative allocation of fishing
effort in the post-implementation period. The largest increase is
observed for the Kiribati EEZ 2 (Kiribati's easternmost EEZ), where the
redistribution of treated vessels caused a 10\% increase in the \emph{relative} allocation of
fishing effort within Kiribati waters. The high seas around Kiribati's EEZ exhibit
similar increases the year of implementation or the year after. Overall, the high seas
receive an incresse in the order of 5\% for 2015.This displacement is likely
causing a crowding effect after PIPA-fishing vessels redistribute to
other areas.

We use the rasterized effort to evaluate if the PIPA closure increased
crowding outside the LSMPA. We count the number of raster cells that
had both treated and control vessels each month (Fig.
\ref{fig:sp_corr}A).

\input{tab/mean_change}

\section{Discussion}\label{discussion}

\section{Conclusion}\label{conclude}

\section{References}\label{references}

\begin{figure}
\centering
\includegraphics{img/redist_trend.pdf}
\caption{\label{fig:redist_trend}Monthly
relative allocation of fishing effort by PIPA-fishing vessels before and
after PIPA by region. Colors indicate the pre- and post- periods. Solid line
across points shows local mean with a region-specific loess smoother.
Vertical dashed lines indicates dates when satellites were added, solid vertical
line indicates PIPA closure. Note that inflection points of
the loess smoother are not affected by the addition of satellites.}
\end{figure}

\begin{figure}
\centering
\includegraphics{img/mean_change.pdf}
\caption{\label{fig:mean_change}Coefficient
estimates for the redistribution regression. Each panel shows the
region-specific coefficients, with 95\% condifence intervals
as error bars. The horizontal dashed line represents 0 change
relative to the 2012 region-specific levels.}
\end{figure}

\begin{figure}
\centering
	\includegraphics{img/PS_VDS_KIR_by_year.pdf}
	\caption{\label{fig:PS_VDS_KIR_by_year}Vessel days spent inside Kiribati waters by group of vessels.}
\end{figure}

\begin{figure}[h]
	\includegraphics[width=0.8\textwidth]{img/Color_V1.png}
	\caption{Map of the Four Protected Areas}
	\label{fig:maps}
\end{figure}

@@

\subsection{Background on Commercial Fishing in Palau}\label{Palau_back}

@@

\clearpage
\subsection{Excluding all USA and TWN vessels}

%\begin{figure}
%\centering
%\includegraphics{Manuscript_files/figure-latex/unnamed-chunk-22-1.pdf}
%\caption{\label{fig:unnamed-chunk-22}\label{fig:q3}Interaction of quarters
%and treatment. Control group excludes Chinese vessels.}
%\end{figure}
%
%\begin{figure}
%\centering
%\includegraphics{Manuscript_files/figure-latex/unnamed-chunk-23-1.pdf}
%\caption{\label{fig:unnamed-chunk-23}\label{fig:q4}Interaction of quarters
%and treatment. Control group is japanese vessels.}
%\end{figure}
%
%\hypertarget{year-month-did-interactions}{%
%\subsubsection{Year-month DID
%interactions}\label{year-month-did-interactions}}
%
%\begin{figure}
%\centering

>>>>

% Keywords are not mandatory, but authors are strongly encouraged to provide them. If provided, please include two to five keywords, separated by the pipe symbol, e.g:
\keywords{Marine Spatial Planning $|$ Fisheries $|$ Marine Conservation}

\begin{abstract}
The Parties to the Nauru Agreement (PNA) is a system for managing the world’s largest skipjack tuna fishery, covering an area of 14.6 million km\textsuperscript{2} in the Pacific Ocean. In 2015, PNA member Kiribati implemented one of the world's largest conservation areas: The Phoenix Island Protected Area (PIPA, 397,447 km\textsuperscript{2}). We use identification of fishing activity via Automatic Identification Systems and causal inference techniques to evaluate the behavioral changes and spatial redistribution of tuna purse seiners due to the implementation of a PIPA. Our work provides three main findings: 1) we observe a crowding effect for the first months after implementation, 2) vessels fish with similar intensity after the implementation, 3) vessels that fished inside the protected area redistribute to adjacent waters, including the high seas, and 4) the redistribution of vessels results in a reduction of ~2500 vessel-days in Kiribati, which represents a loss of ~30 million USD. We use our results to inform predictions of the impacts of a proposed LSMPA in Palau (a PNA member) and estimate potential losses to range from \$2.5 to \$11 million annually. As countries continue to implement LSMPAs as a way to reach ocean protection targets, managers should consider how fishing effort will change in space and time to ensure that fishing effort is not just displaced elsewhere.
\end{abstract}

@@

\dropcap{H}umans are increasingly utilizing the oceans. Multiple ocean uses such as off-shore aquaculture, conservation, energy harvesting, deep-sea mining, and fisheries are likely to compete for space. As we move forward with blue growth, we must understand the potential effects of activities displacing each other and etablish causal links between past management interventions and their outcomes \citep{burgess_2018}.

Global international goals aim to protect 10\% of the ocean environment by 2020. In an effort to meet this target, there has been a rapid increase in MPA coverage \citep{wood_2008,sala_2018}, largely driven by a small number of Large-Scale Marine Protected Areas (LSMPAs), some of which ocurr in the world's largest fishing grounds for tuna. While today a small number of LSMPAs make up at least 80\% of the managed areas in the ocean \citep{toonen_2013}, very little is known about their human dimensions and implication for fisheries \citep{gray_2017}. One issue of particular importance is that of the displacement or redistribution of fishing effort, which may influence the outcomes of a spatial closure \citep{smith_2003}.

LSMPAs were initially assumed to have little social implications due to their general remoteness. However, there have been calls to incorporate the human dimensions into LSMPA management and evaluation \citep{agardy_2011,gray_2017}. Most research incorporating these dimensions has focused on governance and enforcement of LSMPAs \cite{alger_2017,christie_2017}, but they are yet to be the focus of economic analyses \citep{gray_2017}. For example, the anticipation of LSMPAs may lead to preemptive overfishing, which will likely erodes or delay the expected benefits of an intervention \cite{mcdermott_2018}.

One of the most notable Large-Scale Marine Protected Areas recently implemented is the Phoenix Island Protected Area (PIPA) in Kiribati. The area implemented in January 1st of 2015 has an extension of 397,447 km\textsuperscript{2} and represents was implemented within an area that produces 50\% of the World's tuna. Tuna purse seine fisheries in the region are collectively managed under a Vessel-day scheme (VDS) by nine countries commonly refered to as the Parties to the Nauru Agreement (PNA). Members include the Federated States of Micronesia, Kiribati, the Marshall Islands, Nauru, Palau, Papua New Guinea, the Solomon Islands, and Tuvalu. Tokelau joined the PNA group in 2012 and started selling access rights in 2013. The Nauru Agreement regulates access of foreign vessels (\emph{i.e.} those from non-PNA countries). Holding 80\% (14.6 million km\textsuperscript{2}) of historical skipjack tuna purse seine grounds within their Exclusive Economic Zones (EEZ), PNA countries have achieved greater bargaining power when providing fishing access to foreign fleets \cite{havice_2010}. The revenue from access fees represents up to 50\% of government revenue from some of this countries. While no studies have assessed the immplications of PIPA, other PNA members have pledged the implementation of LSMPAs by 2020.

@@

\begin{figure}[H]
\centering
\includegraphics{img/sp_corr.pdf}
\caption{\label{fig:sp_corr}Number of cells that had treated and control vessels (A) and spatial correlation in the presence-absence of treated and control vessels per cell (B). The solid lines represent the 4\textsuperscript{th} degree polynomial fit reported in \ref{tab:sp_corr}}
\end{figure}

@@

The crowding effect combined with the reduction of hours spent in Kiribati and PNA waters overall suggests that treated vessels have redistributed elsewhere, meaning that they no longer buy access fees from PNA countries. To quantify the potential impacts of this leakage, we estimate the total annual vessel-days received by all PNA countries (Fig. \ref{fig:all_PS_VDS_year}). In this case we look at all 313 vessels, but continue to group them as treated and control vessels, as well as other vessels that we do not observe continuously or came into the fishery after the implementation of PIPA. We term this last group as ``others''. We find that the treated vessels spent 2,310 less days in PNA waters when comparing 2015 and 2016, while control and other vessels increased their activities in PNA waters by 5,489 and 4,447 days for the same period, respectively (Figs. \ref{fig:all_PS_VDS_year}A and \ref{fig:included_PS_VDS_year_DiD}). The 2,310 vessel-day reduction by treated vessels between 2015 and 2017 corresponds to \$27.7 million USD at an average price of \$12,000 USD per vessel-day. Looking at the total anual vessel-days allocated by all vessels to all countries, we see that the largest reductions ocurr for Kiribati, while Papua New Guinea exhibits a proportional increase (Fig. \ref{fig:all_PS_VDS_cty_year}). Furthermore, in 2017 treated vessels exhibit the lowest vessel-day allcoation to PNA waters at just 10,026 vessel-days. This is 2,526 vessel-days lower (~\$30.3 million USD) than pre-implementation levels of 2012 - 2013 (mean $\pm$ sd of 12,552.46 $\pm$ 1,396.46; we exclude 2014 due to the blue paradox effect \cite{mcdermott_2018}). % Do you think we need to come up with an identification strategy for this?. A simple DID doesn't work, because the post-trend would assume that there is no upper bound to effort (current DiD estimate of these data suggests a 8,700 vessel-days reduction; or 7,400 vessel-days if we account for the blue paradox with some fixed effects).

\begin{figure}[ht]
\centering
	\includegraphics{img/included_PS_VDS_year_DiD.pdf}
	\caption{\label{fig:included_PS_VDS_year_DiD}Vessel days spent inside A) PNA waters and B) Kiribati waters by group of vessels.}
\end{figure}

We compliment our analysis of change in vessel-days by looking at country-level data. Specifically, we use data compiled by the Pacific Islands Forum Fisheries Agency (FFA\footnote{https://www.ffa.int/node/2050}) where annual revenues from license fees are reported for each country (2008 - 2016; Fig. \ref{fig:revenues}). We find that by 2016, Kiribati's revenue decreased from \$148.8 million USD in 2015 to \$118.3 million USD in 2016, representing a change in 30.5 million USD. We use a local polynomial regression fit to obtain a trend line for the total (\emph{i.e.} summing across countries) revenues between 2008 and 2015, and extrapolate the trend to obtain an expected value for 2016. Our model suggests that revenue should be in the order of \$463.5 million USD, but we observe only \$423.5 million, suggesting that potential losses to the entire PNA are in the order of \$40 million USD.

Catches for each country's EEZ for the 1997 - 2016 period were also obtained from the FFA (Fig. \ref{fig:catches}). Catches in Kiribati waters decreased from 24,051 to 12,894 tonnes between 2015 and 2016 (46.3\% decrease). Similar decreases were observed for The Federal States of Micronesia (60.9\%), Papua New Guinea (43.4\%) and the Solomon Islands (58.5\%). In contrast, Tokelau (due south of Kiribati) showed a 22.3\% increase in catches over the same period.

\begin{figure}[ht]
\centering
	\includegraphics{img/revenues.pdf}
	\caption{\label{fig:revenues}Financial indicators for PNA countries. A) Revenue from fishing licenses fees, B) Total annual revenues for all 9 PNA countries with dashed line showing the local polynomial regression fit to the 2008 - 2015 data and the red dot representing the prediction from this regression. Vertical dashed line in both plots denotes implementation of PIPA.}
\end{figure}

\subsection{Potential Revenue Loss for Palau}

@@

\section{Discussion}

\section{Conclusions}

\section{References}

\onecolumn

\section{Supplementary tables and figures}

@@

\begin{figure}
\centering
\includegraphics{img/all_panels.pdf}
\caption{\label{fig:all_panels}Time series showing monthly averages for eight variables of interest: A) Fishing hours, B) Non-fishing hours at-sea, C) Proportion of fishing hours to total hours at-sea, D) Distance traveled, E) Mean distance from port for fishing evetns, F) Mean distance from shore for fishing events, I) Monthly hours spent in Kiribati waters, J) Proportion of fishing hours spent in PNA waters.}
\end{figure}

\begin{figure}
\centering
	\includegraphics{img/all_PS_VDS_year.pdf}
	\caption{\label{fig:all_PS_VDS_year}Observed vessel-days for all countries by treated, control, and excluded vessels.}
\end{figure}

\begin{figure}
\centering
	\includegraphics{img/all_PS_VDS_cty_year.pdf}
	\caption{\label{fig:all_PS_VDS_cty_year}Observed vessel days by country and year}
\end{figure}

\begin{figure}
\centering
	\includegraphics{img/catches.pdf}
	\caption{\label{fig:catches}Financial indicators for PNA countries. A) Annual catches by EEZ and, B) Annual value of catches by EEZ. Vertical dashed line in both plots denotes implementation of PIPA.}
\end{figure}

\begin{figure}
	\includegraphics[width=0.8\textwidth]{img/Color_V1.png}
	\caption{Map of the Four Protected Areas}
	\label{fig:maps}
\end{figure}

\subsection{Excluding all USA and TWN vessels}

\section{Additional methods}

\subsection{Data}

Automatic Identification Systems (AIS) are on-board devices that provide at-sea safety and prevent ship collisions by broadcasting vessel position, course, and activity to surrounding vessels. These broadcast messages can be received by satellites and land-based antennas. GFW then uses machine learning algorithms (convolutional neural networks) on the broadcast messages to infer what type of fishing is taking place and where it is taking place, thus allowing the estimation of near real-time fishing events globally \citep{kroodsma_2018}.

% All numbers in the following paragraph need to be updated
Our analysis focuses on tune purse seine vessels, the most important fishery for PNA countries. We identify a total of 186 purse seiners that fished in PNA waters at least once before 2015. Vessels that entered the fishery after 2015 were not exposed to the policy intervention in the pre-treatment period and are therefore excluded from our analyses, leaving us with 125 vessels. We identify 90 vessels that have fished inside PIPA at least once since 2012; 72 did so before the announcement of the closure (\emph{i.e.} 09/01/2014 \emph{sensu} \citep{mcdermott_2018}). From these, we only observe before and after data for 64 tuna purse seiners.
On the other hand, 45 vessels never fished inside PIPA during the 2012 - 2015 period, but we only observe befor and after data for 28 of these.

\subsection{Analyses}

\subsubsection{crowding effect}

@@

\subsubsection{Behavioral changes}

We then focus on identifying the response of vessels to the PIPA closure. We use daily fishing and non-fishing hours, daily proportion of fishing vs. non-fishing hours, daily distance traveled (km), distance from shore (km) and distance from home port(km) for fishing and non-fishing events, and proportion of total fishing hours allocated to Kiribati waters and PNA waters as our main outcomes of interest. We compare these outcomes before and after the implementation of PIPA using a Difference-in-Differences approach. Our main specification is the following:

Where $log(y_{i,t})$ is the log-transformed outcome of interest for vessel $i$ on day $t$. A dummy variable $Post_t$ takes the value of 0 for all dates prior to PIPA implementation and a value of 1 for all dates following PIPA implementation. $Treat_i$ is a dummy variable indicating whether a vessel belongs to the treatment ($Treat_i = 1$) or control ($Treat_i = 0$) group. $\alpha$ is the standard intercept term, $beta_1$ captures the temporal trend, $beta_2$ captures the initial difference between treated and control groups, and $beta_3$ is our parameter of interest: the Difference-in-Differences estimate capturing the treatment effect. Finally, $phi_t$ and $\gamma_i$ represent month and flag dummies that account for seasonality or country-level management interventions. We test for other more complex specifications that interact a quarterly dummy or year-month dummies with the treatment group and find qualitatively the same results\footnote{I actually need more time to run these, but I don't think they'll change}.

We discretize spatial units by creating a polygon for PIPA and distinct spatial units for each geographically separate EEZ of each country. Some vessels might shift from EEZs into the high seas, but we are interested in knowing \emph{where} in the high seas, so we incorporate additional regions by using a 5-degree buffer of the high seas. The rest of the high seas are merged into a single spatial unit. For example, if we were to do this only for Kiribati, we would have 8 spatial units: PIPA, three EEZ units, three 1-degree buffers of high seas  around each of the EEZ regions. around each EEZ, and the rest of the high seas. We clipped any overlapping high seas buffers to avoid duplication.

\section{Old Text}

@@

\subsection{Nauru agreement and the Phoenix Islands Protected Area}

@@

\subsection{Palau}

\subsubsection{Palau National Monument}

@@

\subsubsection{Background on Commercial Fishing in Palau}

@@
